---
title: "Covid Draft 3"
author: "Carilli"
date: "12/30/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, options(scipen = 999, digits = 3))
library(tidyverse)
library(gt)
```

# Data
```{r data}
df <- 
  here::here("data", "covid_data_final.csv") %>% 
  read_csv()
```

# Setup

```{r social-capital-models-setup}
models <- list()

model_names <- c(
  "state",
  "family_unity",
  "family_interaction",
  "social_support",
  "community_health",
  "institutional_health",
  "collective_efficacy",
  "philanthropic_health",
  "trust_neighbor",
  "penn_state",
  "putnam_index",
  "ferrara",
  "family_prosperity"
)

social_capital <- c(
  "state_level_dummy",
  "family_unity_dummy",
  "family_interaction_dummy",
  "social_support_dummy",
  "community_health_dummy",
  "institutional_health_dummy",
  "collective_efficacy_dummy",
  "philanthropic_health_dummy",
  "neighbor_dummy",
  "putnam_index",
  "scp_version_of_penn_state_index",
  "alesina_la_ferrara_social_capital_group",
  "family_prosperity_index"
)

controls <- c(
  "median_household_income",
  "percent_adults_with_ba",
  "percent_in_fair_or_poor_health",
  "percent_black",
  "density",
  "percent_rural",
  "percent_65",
  "average_temperature",
  "governor_political_affiliation"
)

```

# Two-way {.tabset}
## Two-Way linear s.i. {.tabset}

### models

```{r two-ways-linear-stringency}
for (i in seq_along(social_capital)) {
  models[[i]] <-
    glue::glue(
      'cases_new ~ stringency_index_lag * ',
      social_capital[i], 
    )
}

model_frame <- tibble(model = models) %>% 
  mutate(model_name = names(model))

model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

estimated_models_twoways_linear <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df, effect = "twoways", index = "state", model = "within")))

models_tidy_linear <- 
  estimated_models_twoways_linear %>% 
  transmute(model_names, betas = purrr::map(fit, broom::tidy))

models_glanced_linear <- 
  estimated_models_twoways_linear %>% 
  transmute(model_names, stats = purrr::map(fit, broom::glance))
```

### summary
```{r linear-summary}
estimated_models_twoways_linear$fit %>% 
  map(summary)
```


### mar-may
```{r linear-mar-may}
estimated_models_twoways_linear_mar_may <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date <= "2020-03-01"), 
                                    effect = "twoways", index = "state", model = "within")))
estimated_models_twoways_linear_mar_may$fit %>% 
  map(summary)
```

### jun-aug

```{r linear-june-aug}
estimated_models_twoways_linear_jun_aug <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-06-01" & date <= "2020-08-31"), 
                                    effect = "twoways", index = "state", model = "within")))
estimated_models_twoways_linear_jun_aug$fit %>% 
  map(summary)
```

### sep-nov

```{r linear-sep-nov}
estimated_models_twoways_linear_sep_nov <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-09-01"), 
                                    effect = "twoways", index = "state", model = "within")))
estimated_models_twoways_linear_sep_nov$fit %>% 
  map(summary)
```


## Tables {.tabset}
### table 1a
```{r table-1a, eval = FALSE}
models_tidy_linear %>% 
  left_join(models_glanced_linear, by = "model_names") %>% 
  unnest(c(betas, stats), names_repair = "universal") %>% 
  select(model_names, estimate, std.error, r.squared, statistic...9, p.value...10, nobs) %>% 
  rename(f_stat = statistic...9, p_val = p.value...10) %>% 
  mutate(term = rep(c("si", "interact"), times = 13), .before = estimate) %>% 
  mutate(across(where(is.numeric), round, 3)) %>%
  mutate(nobs = as.integer(nobs)) %>% 
  pivot_longer(-c(model_names, term),
               names_to = "measure",
               values_to = "value") %>% 
  pivot_wider(names_from = model_names) %>% 
  slice(1, 2, 7:12) %>% 
  mutate(measure = c(
    "Stringency Index",
    "se",
    "Stringency Index * Social Capital Index",
    "se",
    "R-sqr",
    "F",
    "p-value",
    "N"
  )
  ) %>% 
  # arrange(match(measure, c("estimate", "std.error")), term) %>% 
  gt() %>% 
  fmt_number(model_names, rows = 8, decimals = 0) %>% 
  cols_label(measure = "",
             state = "Total Social Capital",
             family_unity = "Family Unity",
             family_interaction = "Family Interaction",
             social_support = "Social Support",
             community_health = "Community Health",
             institutional_health = "Instutional Health",
             collective_efficacy = "Collective Efficacy",
             philanthropic_health = "Philanthropic Health",
             trust_neighbor = "Trust Neighbors"
  ) %>% 
  tab_spanner("Sub-Index", model_names[2:9]) %>% 
  cols_align("center") %>% 
  cols_hide(c(1, 12:15)) %>% 
  opt_table_lines("default") %>% 
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_body(
      columns = vars(measure),
      rows = 7:8
    )
  ) %>% 
  tab_footnote(
    footnote = "Siginicant at the 1% level",
    locations = cells_body(
      columns = vars(model_names),
      rows = c(1,3)
    )
  ) %>% 
  opt_footnote_marks(marks = c("*", "+")) %>% 
  tab_header(
    title = "Two-way Fixed Effects Results",
    subtitle = "Joint Economic Committee Social Capital Indicies")
```

### table 1b
```{r table-1b, eval = FALSE}
models_tidy_linear %>% 
  left_join(models_glanced_linear, by = "model_names") %>% 
  unnest(c(betas, stats), names_repair = "universal") %>% 
  select(model_names, estimate, std.error, r.squared, statistic...9, p.value...10, nobs) %>% 
  rename(f_stat = statistic...9, p_val = p.value...10) %>% 
  mutate(term = rep(c("si", "interact"), times = 13), .before = estimate) %>% 
  mutate(across(where(is.numeric), round, 3)) %>%
  mutate(nobs = as.integer(nobs)) %>% 
  pivot_longer(-c(model_names, term),
               names_to = "measure",
               values_to = "value") %>% 
  pivot_wider(names_from = model_names) %>% 
  slice(1, 2, 7:12) %>% 
  mutate(measure = c(
    "Stringency Index",
    "se",
    "Stringency Index * Social Capital Index",
    "se",
    "R-sqr",
    "F",
    "p-value",
    "N"
  )
  ) %>% 
  # arrange(match(measure, c("estimate", "std.error")), term) %>% 
  gt() %>% 
  fmt_number(model_names, rows = 8, decimals = 0) %>% 
  cols_label(measure = "",
             penn_state = "Penn State Index (SCP)",
             putnam_index = "Putnam Index",
             ferrara = "Alesina La Ferrar",
             family_prosperity = "Family Prosperity"
  ) %>% 
  # tab_spanner("Sub-Index", model_names[2:9]) %>% 
  cols_align("center") %>% 
  cols_hide(c(1, 3:11)) %>% 
  opt_table_lines("default") %>% 
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_body(
      columns = vars(measure),
      rows = 7:8
    )
  ) %>% 
  tab_footnote(
    footnote = "Siginicant at the 1% level",
    locations = cells_body(
      columns = vars(model_names),
      rows = c(1,3)
    )
  ) %>% 
  opt_footnote_marks(marks = c("*", "+")) %>% 
  tab_header(
    title = "Two-way Fixed Effects Results",
    subtitle = "Social Capital Indices")
```


# Two-way quadratic si {.tabset}
## models

```{r two-ways-quadratic-stringency}
for (i in seq_along(social_capital)) {
  models[[i]] <-
    glue::glue(
      'cases_new ~ stringency_index_lag * ',
      social_capital[i], 
      ' + I(stringency_index_lag^2)'
    )
}

model_frame <- tibble(model = models) %>% 
  mutate(model_name = names(model))

model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

estimated_models_twoways_quadratic <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df, effect = "twoways", index = "state", model = "within")))

models_tidy_quadratic <- 
  estimated_models_twoways_quadratic %>% 
  transmute(model_names, betas = purrr::map(fit, broom::tidy))

models_glanced_quadratic <- 
  estimated_models_twoways_quadratic %>% 
  transmute(model_names, stats = purrr::map(fit, broom::glance))
```


## summary
```{r quadratic-summary}
estimated_models_twoways_quadratic$fit %>% 
  map(summary)
```


## mar-may
```{r quadratic-mar-may}
estimated_models_twoways_quadratic_mar_may <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date <= "2020-03-01"), 
                                    effect = "twoways", index = "state", model = "within")))
estimated_models_twoways_quadratic_mar_may$fit %>% 
  map(summary)
```

## jun-aug

```{r quadratic-june-aug}
estimated_models_twoways_quadratic_jun_aug <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-06-01" & date <= "2020-08-31"), 
                                    effect = "twoways", index = "state", model = "within")))
estimated_models_twoways_quadratic_jun_aug$fit %>% 
  map(summary)
```

## sep-nov

```{r quadratic-sep-nov}
estimated_models_twoways_quadratic_sep_nov <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-09-01"), 
                                     effect = "twoways", index = "state", model = "within")))
estimated_models_twoways_quadratic_sep_nov$fit %>% 
  map(summary)
```

## tables
### table 2a
```{r table-2a, eval=FALSE}
models_tidy_quadratic %>% 
  left_join(models_glanced_quadratic, by = "model_names") %>% 
  unnest(c(betas, stats), names_repair = "universal") %>% 
  select(model_names, estimate, std.error, r.squared, statistic...9, p.value...10, nobs) %>% 
  rename(f_stat = statistic...9, p_val = p.value...10) %>% 
  mutate(term = rep(c("si", "interact", "quadratic"), times = 13), .before = estimate) %>% 
  mutate(across(where(is.numeric), round, 3)) %>%
  mutate(nobs = as.integer(nobs)) %>% 
  pivot_longer(-c(model_names, term),
               names_to = "measure",
               values_to = "value") %>% 
  pivot_wider(names_from = model_names)
```





# One-way Linear No Controls{.tabset}
## One-way linear
```{r one-way-linear}
for (i in seq_along(social_capital)) {
  models[[i]] <-
    glue::glue(
      'cases_new ~ stringency_index_lag * ',
      social_capital[i]
    )
}

model_frame <- tibble(model = models) %>% 
  mutate(model_name = names(model))

model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

estimated_models_oneway_linear <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df, effect = "time", index = "state", model = "within")))

estimated_models_oneway_linear$fit %>% 
  map(summary)

```


## mar - may
```{r one-way-linear-mar-may}
for (i in seq_along(social_capital)) {
  models[[i]] <-
    glue::glue(
      'cases_new ~ stringency_index_lag * ',
      social_capital[i]
    )
}

model_frame <- tibble(model = models) %>% 
  mutate(model_name = names(model))

model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

estimated_models_oneway_linear_mar_may <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date <= "2020-05-31"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_linear_mar_may$fit %>% 
  map(summary)

```

## jun-aug

```{r one-way-linear-jun-aug}
for (i in seq_along(social_capital)) {
  models[[i]] <-
    glue::glue(
      'cases_new ~ stringency_index_lag * ',
      social_capital[i]
    )
}

model_frame <- tibble(model = models) %>% 
  mutate(model_name = names(model))

model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

estimated_models_oneway_linear_jun_aug <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-06-01" & date <= "2020-08-31"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_linear_jun_aug$fit %>% 
  map(summary)

```
## sep - nov


```{r one-way-linear-sep-nov}
for (i in seq_along(social_capital)) {
  models[[i]] <-
    glue::glue(
      'cases_new ~ stringency_index_lag * ',
      social_capital[i]
    )
}

model_frame <- tibble(model = models) %>% 
  mutate(model_name = names(model))

model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

estimated_models_oneway_linear_sep_nov <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-09-01"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_linear_sep_nov$fit %>% 
  map(summary)

```
# One-way Linear with controls {.tabset}
## one-way linear with controls
```{r one-way-linear-controls}
for (i in seq_along(social_capital)) {
  models[[i]] <-
    glue::glue(
      'cases_new ~ stringency_index_lag * ',
      social_capital[i],
      ' + ',
      glue::glue_collapse(controls, " + ")
    )
}

model_frame <- tibble(model = models) %>% 
  mutate(model_name = names(model))

model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

estimated_models_oneway_linear_controls <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df, effect = "time", index = "state", model = "within")))

estimated_models_oneway_linear_controls$fit %>% 
  map(summary)
```
## mar - may
```{r one-way-linear-controls-mar-may}
estimated_models_oneway_linear_controls_mar_may <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date <= "2020-05-31"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_linear_controls_mar_may$fit %>% 
  map(summary)
```

## jun-aug

```{r one-way-linear-controls-jun-aug}
estimated_models_oneway_linear_controls_jun_aug <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-06-01" & date <= "2020-08-31"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_linear_controls_jun_aug$fit %>% 
  map(summary)
```


## sep-nov

```{r one-way-linear-controls-sep-nov}
estimated_models_oneway_linear_controls_sep_nov <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-09-01"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_linear_controls_sep_nov$fit %>% 
  map(summary)
```



# One-way quadratic no controls {.tabset}
## one-way quadratic
```{r one-way-quadratic}
for (i in seq_along(social_capital)) {
  models[[i]] <-
    glue::glue(
      'cases_new ~ stringency_index_lag * ',
      social_capital[i],
      ' + I(stringency_index_lag^2)'
    )
}

model_frame <- tibble(model = models) %>% 
  mutate(model_name = names(model))

model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

estimated_models_oneway_quadratic <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df, effect = "time", index = "state", model = "within")))

estimated_models_oneway_quadratic$fit %>% 
  map(summary)

```

## mar - may

```{r one-way-quadratic-mar-may}
estimated_models_oneway_quadratic_mar_may <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date <= "2020-05-31"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_quadratic_mar_may$fit %>% 
  map(summary)
```

## jun-aug

```{r one-way-quadratic-jun-aug}
estimated_models_oneway_quadratic_jun_aug <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-06-01" & date <= "2020-08-31"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_quadratic_jun_aug$fit %>% 
  map(summary)
```

## sep-nov

```{r one-way-quadratic-sep-nov}
estimated_models_oneway_quadratic_sep_nov <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-09-01"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_quadratic_sep_nov$fit %>% 
  map(summary)
```



# One-way quadratic with controls {.tabset}

## one-way quadratic with controls

```{r one-way-quadratic-controls}
for (i in seq_along(social_capital)) {
  models[[i]] <-
    glue::glue(
      'cases_new ~ stringency_index_lag * ',
      social_capital[i],
      ' + I(stringency_index_lag^2)',
      ' + ', 
      glue::glue_collapse(controls, " + ")
    )
}

model_frame <- tibble(model = models) %>% 
  mutate(model_name = names(model))

model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

estimated_models_oneway_quadratic_contols <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df, effect = "time", index = "state", model = "within")))

estimated_models_oneway_quadratic_contols$fit %>% 
  map(summary)
```

## mar - may

```{r one-way-quadratic-controls-mar-may}
estimated_models_oneway_quadratic_controls_mar_may <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date <= "2020-05-31"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_quadratic_controls_mar_may$fit %>% 
  map(summary)
```

## jun - aug

```{r one-way-quadratic-controls-jun-aug}
estimated_models_oneway_quadratic_controls_jun_aug <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-06-01" & date <= "2020-08-31"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_quadratic_controls_jun_aug$fit %>% 
  map(summary)
```

## sep - nov

```{r one-way-quadratic-controls-sep-nov}
estimated_models_oneway_quadratic_controls_sep_nov <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df %>% filter(date >= "2020-09-01"), 
                                    effect = "time", index = "state", model = "within")))

estimated_models_oneway_quadratic_controls_sep_nov$fit %>% 
  map(summary)
```

